NLTK: The Natural Language Toolkit

### 2002-05-13 technical/acl2002/acl02.tex
Wrote survey section, new paragraph in conclusion
svn/trunk@612

% [EL] ``punitive'' like a strange word.
% [EL] I couldn't find a REF for ``executable pseudocode'' -- closest
% I got was http://www.thinkware.se/cgi-bin/thinki.cgi/PythonQuotes
In surveying the available languages, we believe that Python offers an
especially good fit to the above requirements.  Python is an
object-oriented scripting language developed by Guido van Rossum
and available on all platforms \cite{python}.  Python offers a
shallow learning curve; it was designed to be easily learnt by
children \cite{rossum99}.  As an interpreted language, Python is
suitable for rapid prototyping.  Python code is exceptionally
readable, and it has been praised as ``executable pseudocode.''
Python is an object-oriented language, but not punitively so, and it
is easy to encapsulate data and methods inside Python classes.
Finally, Python has an interface to the Tk graphics toolkit
\cite{tkinter}, and writing graphical interfaces is straightforward.

@@

% (Leaving this out for now..)
%The Language: Python was new to 95\% of students.  Some were initially
%bothered about its block structuring, which is sensitive to white
%space.  All quickly learnt the language, and reported that any initial
%reservations were unfounded.  Many students said that they enjoyed
%writing Python code.

% Where does this go??  Leaving it out for now..
%At the beginning of the semester, we distributed 2 CD-ROMs to the
%students, containing NLTK and all of the corpera that we would use for
%the class.

@@

This section will include a discussion of other approaches to
implementing computational linguistics courseware, including
\emph{Programming for Linguistics: Java Technology for Language
Researchers}\cite{Hammond02} and \emph{the OpenNLP
project}\cite{opennlp}.

@@

>>>>

% [EL] ``punitive'' like a strange word.
% [EL] I couldn't find a REF for ``executable pseudocode'' -- closest
% I got was http://www.thinkware.se/cgi-bin/thinki.cgi/PythonQuotes
% [SB] changed Python citation to URL
In surveying the available languages, we believe that Python offers an
especially good fit to the above requirements.  Python is an
object-oriented scripting language developed by Guido van Rossum
and available on all platforms (\url{www.python.org}).  Python offers a
shallow learning curve; it was designed to be easily learnt by
children \cite{rossum99}.  As an interpreted language, Python is
suitable for rapid prototyping.  Python code is exceptionally
readable, and it has been praised as ``executable pseudocode.''
Python is an object-oriented language, but not punitively so, and it
is easy to encapsulate data and methods inside Python classes.
Finally, Python has an interface to the Tk graphics toolkit
\cite{tkinter}, and writing graphical interfaces is straightforward.

@@

@@

The computational component of computational linguistics courses takes
many forms.  In this section we briefly review a selection of approaches,
classified according to the (original) target audience.

{\bf Linguistics Students.}
Various books introduce programming or computing to linguists.
These are elementary on the computational side, providing a
gentle introduction to students having no prior experience
in computer science.  Examples of such books are:
\emph{Using Computers in Linguistics}
\cite{Lawler98}, and
\emph{Programming for Linguistics: Java Technology for Language
Researchers} \cite{Hammond02}.

{\bf Grammar Developers.}
Infrastructure for grammar development has a long history in
unification-based (or constraint-based) grammar frameworks, from DCG
\cite{PereiraWarren80} to HPSG \cite{PollardSag94}.  Recent work includes
\cite{Copestake00,Baldridge02}.  A concurrent development has been the
finite state toolkits, such as the Xerox toolkit \cite{Beesley02}.  This
work has found widespread pedagogical application.

{\bf Other Researchers and Developers.}
A variety of toolkits have been created for research or R\&D
purposes.  Examples include
the \emph{CMU-Cambridge Statistical Language Modeling Toolkit}
\cite{Clarkson97},
the \emph{EMU Speech Database System} \cite{Harrington99},
the \emph{General Architecture for Text Engineering} \cite{Bontcheva02},
the \emph{Maxent Package for Maximum Entropy Models} \cite{maxent},
and the \emph{Annotation Graph Toolkit} \cite{MaedaBird02}.
Although not originally motivated by pedagogical needs, all of these
toolkits have pedagogical applications and many have already been
used in teaching.

@@

NLTK is unique in its combination of three factors.
First, it was deliberately designed as courseware and gives pedagogical
goals primary status.  Second, its target audience consists of both
linguists and computer scientists, and it is accessible and challenging
at many levels of prior computational skill.  Finally, it is based on
an object-oriented scripting language supporting rapid prototyping and
literate programming.
